---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# btw

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/btw)](https://CRAN.R-project.org/package=btw)
[![R-CMD-check](https://github.com/posit-dev/btw/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/posit-dev/btw/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/posit-dev/btw/graph/badge.svg)](https://app.codecov.io/gh/posit-dev/btw)
<!-- badges: end -->

btw helps you describe your computational environment to LLMs.

* When used **interactively**, `btw()` assembles context on your R environment, package documentation, and working directory, copying the results to your clipboard for easy pasting into chat interfaces.
* The `btw()` function wraps methods that can be easily incorporated into **ellmer tool calls** for describing various kinds of objects in R. To equip your ellmer chat with the ability to peruse documentation, check out the objects in your R environment, and explore your working directory, pass your chat to `btw_register_tools()`.

## Installation

You can install the development version of btw like so:

``` r
pak::pak("posit-dev/btw")
```

## Example

```{r}
library(btw)
```

### Interactive use

The `btw()` function allows you to compile information about all sorts of R stuff and copy it to your clipboard.

* Pass any R object to `btw()`, like `btw(mtcars)`, `btw(dplyr::across)` or `btw(globalenv())`.
* `btw("{dplyr}")` will describe the dplyr package via the function reference and (if available) introductory vignette.
* `btw("?dplyr::across")` or `btw(?dplyr::across)` includes the help pages for `dplyr::across()`.
* `btw("./path")` will read the lines of the file at that path (or list the files at that path if it's a directory).
* `btw("@current_file")` and `btw("@current_selection")` reads the contents of the current editor or selection in RStudio, Positron, or anywhere the [rstudioapi](https://rstudio.github.io/rstudioapi) is supported.
* `btw("@platform_info")` describes your R version, operating system, and locale information.
* `btw()` with `"@attached_packages"`, `"@loaded_packages"`, or `"@installed_packages"` includes a listing of attached (i.e. with `library()`), loaded (in use in the session, often indirectly), or installed packages.

When passed multiple arguments, `btw()` will concatenate each description. For example, you could run:

```{r}
#| eval: false
btw(mtcars, "{btw}", ?btw::btw)
```

```
✔ btw copied to the clipboard!
```

The following would be attached to your clipboard:

```{r}
#| comment: ""
#| echo: false
cli::cat_line(btw(mtcars, "{btw}", ?btw::btw))
```

You can also just call `btw()` with no inputs, which will describe the objects in your global environment.

### Supercharging assistants

`btw_client()` equips LLM chats with all of the same capabilities: peruse
documentation, check out objects in your R environment, and explore your
working directory.

```{r}
#| eval: false
ch <- btw_client()

ch$chat("Hey!")
```

```
Hello! I'm here to help you work with R-related information and data. I can help
you:

1. Look up information about installed R packages
2. Get help documentation for specific R packages and functions
3. Examine and manipulate data frames
4. Check what's available in the current environment

What would you like to know about? Feel free to ask a specific question, and I'll
use the available tools to help you find the information you need.
```

You can also provide `btw_client()` with objects and documentation to include as context in the chat:

```{r}
#| eval: false

ch <- btw_client(mtcars)
ch$chat("Write base R code to summarize average mpg of cars with 4-6 cylinders.")
```
````
I'll help you write base R code to calculate the average MPG for cars with 4-6
cylinders using the mtcars dataset.

Here's the code:

```R
mean(mtcars$mpg[mtcars$cyl >= 4 & mtcars$cyl <= 6])
```

Alternatively, you could also write it with `subset()` from base R:

```R
mean(subset(mtcars, cyl >= 4 & cyl <= 6)$mpg)
```

Both approaches will give you the average MPG for cars with 4-6 cylinders.
Given the data shown in the context, this will include cars with 4 and 6
cylinders (as the `cyl` column only contains values 4, 6, and 8).
````

`btw_client()` uses `ellmer::chat_claude()` by default, or you can customize the chat client used in one of two ways:

* Set the `btw.chat_client` option (possibly in your `.Rprofile` with `usethis::edit_r_profile()`), or
* Provide your own chat object via the `client` argument.

```{r}
#| eval: false

options(btw.chat_client = ellmer::chat_ollama(model = "llama3.1:8b"))
ch <- btw_client()

# same as above
ch <- btw_client(client = ellmer::chat_ollama(model = "llama3.1:8b"))
```

Alternatively, you can call `btw_app()` to jump straight into a Shiny chat app.

For fully customized chat clients, you can use `btw_register_tools()` to add btw
tools to an existing chat interface. Each of the individual tools registered by
`btw_register_tools()` are themselves exported, and can be selected via the
`tools` argument of `btw_register_tools()` (also available in `btw_client()` or
`btw_app()`).
