---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# btw

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/btw)](https://CRAN.R-project.org/package=btw)
[![R-CMD-check](https://github.com/posit-dev/btw/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/posit-dev/btw/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/posit-dev/btw/graph/badge.svg)](https://app.codecov.io/gh/posit-dev/btw)
<!-- badges: end -->

btw helps you describe your computational environment to LLMs.

* When used **interactively**, `btw()` assembles context on your R environment, package documentation, and working directory, copying the results to your clipboard for easy pasting into chat interfaces.
* The `btw()` function wraps methods that can be easily incorporated into **ellmer tool calls** for describing various kinds of objects in R. To equip your ellmer chat with the ability to peruse documentation, check out the objects in your R environment, and explore your working directory, pass your chat to `btw_register_tools()`.

## Installation

You can install the development version of btw like so:

``` r
pak::pak("posit-dev/btw")
```

## Example

```{r}
library(btw)
```

### Interactive use

The `btw()` function allows you to compile information about all sorts of R stuff and copy it to your clipboard.

* Pass any R object to `btw()`, like `btw(mtcars)`, `btw(dplyr::across)` or `btw(globalenv())`. Strings will be evaluated as R code and dispatched to the appropriate method unless they match a shortcut described below.
* `btw("{dplyr}")` will describe the dplyr package via the function reference and (if available) introductory vignette.
* `btw("./path")` will read the lines of the file at that path (or list the files at that path if it's a directory).

When passed multiple arguments, `btw()` will concatenate each description. For example, you could run:

```{r}
#| eval: false
btw(mtcars, "{btw}", btw::btw)
```

```
✔ btw copied to the clipboard!
```

The following would be attached to your clipboard:

```{r}
#| comment: ""
#| echo: false
cli::cat_line(btw(mtcars, "{btw}", btw::btw))
```

You can also just call `btw()` with no inputs, which will describe the objects in your global environment.

### Supercharging assistants

`btw_register_tools()` equips LLM chats with all of the same capabilities: peruse
documentation, check out objects in your R environment, and explore your
working directory. Use it by calling the function on an existing ellmer
chat:

```{r}
#| eval: false
ch <- ellmer::chat_claude()
ch <- btw_register_tools(ch)

ch$chat("Hey!")
```

```
Hello! I'm here to help you work with R-related information and data. I can help
you:

1. Look up information about installed R packages
2. Get help documentation for specific R packages and functions
3. Examine and manipulate data frames
4. Check what's available in the current environment

What would you like to know about? Feel free to ask a specific question, and I'll
use the available tools to help you find the information you need.
```

Each of the individual tools registered by `btw_register_tools()` are themselves exported if you want to pick and choose which ones to use.
