---
title: "Configure btw with btw.md"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configure chats with btw.md}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(btw)
```

A **`btw.md` file** is a Markdown file that saves how btw should set up a chat, so each chat in your project starts already set up: the right model, the right tools, and instructions you'd otherwise repeat.

A `btw.md` file has two parts.
The YAML front matter at the top chooses the model and tools, and the Markdown body below it gives the model instructions about your project.
`btw_app()` and `btw_client()` find the file in your project or home directory and apply it automatically.

If you use other coding agents, you may already keep project instructions in an [`AGENTS.md`](https://agents.md/) or [`CLAUDE.md`](https://code.claude.com/docs/en/memory) file.
A `btw.md` file plays the same role for btw, and btw falls back to those files when a project doesn't have one (see [Where btw looks for `btw.md`](#where-btw-looks)).
Using `btw.md` gives you a few things those files don't:

- **It configures the chat, not just the instructions.** The front matter chooses the model, tools, and options. btw ignores front matter in `CLAUDE.md`, and the other tools that read `AGENTS.md` don't expect front matter there.
- **It has a user-level version.** btw combines a `btw.md` in your home directory, such as `~/.btw/btw.md`, with each project's file. If there's no user-level `btw.md`, btw falls back to a user-level `AGENTS.md` in the cross-tool `~/.agents/` directory. A user-level `CLAUDE.md` is never used.
- **It keeps btw-specific instructions separate.** Instructions about btw's tools or your R workflow don't end up in the files that other coding agents read.

In this vignette, we'll set up a chat by hand, save that setup in a `btw.md` file, and then learn how btw decides which file to use.
Custom agents are written in the same format, so this vignette is also the groundwork for [Custom agents](custom-agents.html).
To follow along, you'll need btw, an [ellmer](https://ellmer.tidyverse.org/) chat provider you can use, and [shinychat](https://posit-dev.github.io/shinychat/) to run the chat app.

## Start with a chat you set up by hand

`btw_app()` starts a chat app in your R session.
Its `client` argument takes an ellmer chat object, which chooses the provider and model.
This example uses `ellmer::chat_posit()`; substitute the `chat_*()` function for the provider you use.

```{r, eval = FALSE}
library(btw)
btw_app(client = ellmer::chat_posit())
```

Inside the app, you can customize which tools the model can use in the chat.
The tool menu lists all available tools, and you can turn each one on or off.
Similarly, you can also edit the system prompt (the set of instructions sent to the model at the start of a chat).
Together, the model, the tools, and the system prompt make up the chat's configuration.

Those choices last only as long as the app is open, but you can use a `btw.md` file to save those choices so that every btw chat in the project starts from the same place.

## Save the setup in `btw.md`

Suppose you're working on a research project and want the chat to read your files, look up R documentation, check installed packages, and run R code.
Create a file named `btw.md` at the root of the project with this content:

``` text
---
client: posit
tools: [files, docs, sessioninfo, run]
---

You are a research data science assistant.
Inspect the project files before answering.
```

The two `---` lines mark the start and end of the front matter.
The front matter configures the chat: `client: posit` chooses the provider, the same way `ellmer::chat_posit()` did above, and `tools` chooses which btw tools the chat can use.
The body gives the model instructions to steer its behavior.
In short, the YAML configures btw and the markdown configures the model.

Once you've created `btw.md`, you can start `btw_app()` without any arguments from the project directory:

```{r, eval = FALSE}
btw_app()
```

btw finds `btw.md`, creates the client and the tools it describes, and adds the body to the system prompt.
If you open the system prompt in the app, you'll find the body under a "Project Context" heading, after btw's own instructions about your R session and tools.
Because the setup now lives in a file, you can read it, edit it, commit it, and share it with collaborators.

## Create and edit the file

You can write `btw.md` by hand, or let btw start it for you.

`use_btw_md()` creates a `btw.md` file from a template at the project root, and in an R package it also adds the file to `.Rbuildignore`.
`use_btw_md("user")` creates the user-level file.
`edit_btw_md()` opens the file that `btw_client()` would use, so you can check which one is in effect.

```{r, eval = FALSE}
use_btw_md()
edit_btw_md()
```

To draft the instructions for an existing project, `btw_task_create_btw_md()` starts a chat in which the model explores the project and writes a `btw.md` with you.

The rest of this vignette covers each part of the file in more detail.

## Choose the model with `client`

The `client` field takes the same `provider` or `provider/model` strings as `ellmer::chat()`.
`client: posit` uses the provider's default model; add a model name after a slash to choose a specific one:

``` text
---
client: posit/claude-sonnet-5-5
---
```

For more control, write `client` as a mapping.
`provider` chooses the `ellmer::chat_*()` function, `model` becomes its `model` argument, and any other keys are passed to that function as arguments.
This example points `ellmer::chat_ollama()` at a server on another machine:

``` text
---
client:
  provider: ollama
  model: "gpt-oss:20b"
  base_url: "http://my-company.example.com:11434"
---
```

You can also list several clients.
When you start `btw_client()` interactively, btw shows a menu of the choices; in non-interactive code, it uses the first one.
If you name the entries, you can pass a name to the `client` argument, as in `btw_client(client = "local")`:

``` text
---
client:
  sonnet: posit/claude-sonnet-5-5
  local:
    provider: ollama
    model: "gpt-oss:20b"
---
```

```{r, eval = FALSE}
btw_app()
#>
#> ── Select a client ──
#>
#> 1. sonnet: posit/claude-sonnet-5-5 [default]
#> 2. local: ollama/gpt-oss:20b
#>
#> Enter choice [1-2, alias name, or 0 to exit]: sonnet
```

The `client` field is a default, not a requirement.
A `client` argument to `btw_app()` or `btw_client()` takes priority over it, and so does the `btw.client` R option.

## Choose tools with `tools`

The `tools` field lists tool groups, individual tool names, or both.
A group name such as `files` includes every tool in that group.
To see which tools a group contains, pass the same names to `btw_tools()`:

```{r}
names(btw_tools("files"))
```

The `files` group includes tools that write and edit files, not only ones that read them.
When the chat only needs to read, name the individual tools instead.
You can drop the `btw_tool_` prefix from tool names in `btw.md`:

``` text
---
tools: [docs, files_read, files_search, files_list]
---
```

If you leave `tools` out, the chat gets btw's default set of tools.
The R code tool in the `run` group isn't part of that default, because it runs code in your R session without asking you first.
Naming `run` in `tools` turns it on for that chat; see `?btw_tool_run_r` before you do.

As with `client`, a `tools` argument to `btw_app()` or `btw_client()`, or the `btw.tools` R option, takes priority over the field in `btw.md`.

## Set btw options with `options`

Some btw settings are R options whose names start with `btw.`, and the `options` field sets them for chats in this project.
Write each option name without the `btw.` prefix, nesting one level for each remaining part of the name.
For example, `btw.subagent.client` chooses the model for **subagents**, the separate chat sessions that your chat can start through a tool call:

``` text
---
client: posit/claude-sonnet-5-5
options:
  subagent:
    client: posit/zai-org/GLM-5.3-Flash
---
```

With this file, the main chat uses Claude Sonnet and its subagents use the smaller, faster GLM 5.3 Flash model.
You can learn more about custom agents in `vignette("custom-agents")`.
Because options are tool-specific, they are listed in each tool's help page, such as `?btw_tool_agent_subagent`.

## Write instructions in the body

Everything after the front matter becomes part of the system prompt.
Use the body for what you'd otherwise repeat at the start of every conversation: your coding style, terms that mean something specific in your field, the goals of the project, or where to find important files.
Write it in whatever structure suits the project; btw doesn't require particular headings.

Sometimes the file holds notes that you want to keep but don't want the model to read, such as finished tasks in a project plan.
Wrap those lines in `<!-- HIDE -->` and `<!-- /HIDE -->` comments:

``` text
## Current task

Add a sensitivity analysis to `analysis.qmd`.

<!-- HIDE -->
## Done

- Cleaned the survey data.
<!-- /HIDE -->
```

btw removes the hidden lines before adding the body to the system prompt.
A `<!-- HIDE -->` without a closing `<!-- /HIDE -->` hides everything to the end of the file.

## Where btw looks for `btw.md` {#where-btw-looks}

When you start a chat, btw looks for a project file first.
It starts in the working directory and moves up through its parents, stopping at the project root, which btw recognizes by a file such as `DESCRIPTION`, `.git`, or an `.Rproj` file.
btw uses the closest `btw.md` it finds.
If there isn't one, it uses an [`AGENTS.md`](https://agents.md/) file, and then a `CLAUDE.md` file.
btw reads the front matter in `AGENTS.md` like `btw.md`, but ignores the front matter in `CLAUDE.md`.
btw reads only one of these files, so if your project has both `btw.md` and `AGENTS.md`, btw uses `btw.md` and ignores `AGENTS.md`.

btw also reads a **user-level** `btw.md` that applies to all of your projects, such as `~/.btw/btw.md`.
If there's no user-level `btw.md`, btw falls back to a user-level `AGENTS.md` in the cross-tool `~/.agents/` directory.
The two files are combined:

- The project's `client` and `tools` fields are used when they're present; otherwise the user-level fields are used.
- The `options` from both files are combined, and the project's value wins when both set the same option.
- Both bodies are included in the system prompt, the user-level instructions first.

This lets you keep your general preferences, such as a default model or coding style, in the user-level file and keep project details in the project file.
See `?btw-config` for every location btw checks.

To use a particular file instead of searching, pass its path to `path_btw`.
To start a chat without any `btw.md` file, use `path_btw = FALSE`.
Or pass a named list to control project and user context separately, e.g. `path_btw = list(user = FALSE)`; the `btw.client.path_btw_user` option sets the default for the user scope.

```{r, eval = FALSE}
btw_app(path_btw = "analysis/btw.md")

btw_client(path_btw = FALSE)

btw_client(path_btw = list(user = FALSE))
```

## Next: the same format for agents

You now have a `btw.md` file that chooses the model, the tools, and the instructions for the chat you talk to, and you know how btw finds and combines these files.

A custom agent file uses the same format.
Add a `name` and a `description` to the front matter and save the file in `.btw/agents/`, and btw turns it into a tool your chat can call.
Each call starts a subagent session configured by that file.
`vignette("custom-agents")` builds three of these agents, starting from the `btw.md` file in this vignette.
